---
title: 'Analyses for the paper: Measuring emotions during learning'
author: "Franziska Hirt"
date: "23 Mai 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = T, warning = F, message = F)
```

---
title: "R Notebook"
author: "Franziska Hirt"
output:
  pdf_document: default
---

# PREPARATIONS
```{r}
# load packages
library(tidyverse)
library(stats)
library(scatr)
library(jmv)
library(rstan)
library(brms)
library(bayesplot)
library(ggmcmc) # for ggs posterior plot
library(ggthemes)
library(papaja)

# set ggplot theme
theme_set(theme_default())

# set rstan options
rstan::rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

## Selection of statistical methods
- Bayesian statistics as more intuitive to interprete

- Mixed models, because events (level 1) nested within subjects (level 2). But only random intercepts included, as random effects when only two rows per participant are hardly identifyable.

- Treat emotional self-reports as ordinal outcome variables including edge effects in the data (i.e. many data points in the highest/lowest category). Cummulative family as we understand the Likert-scales as the categorization of a latent continuous construct (cf. Bürkner & Vuorre, 2019)

- Assummption of equal variances: "If unequal variances are theoretically possible -- and they usually are -- we also recommend incorporating them into the model" (Burkner & Vuorre, 2019). However, models allowing for unequal variances did not converge and were therefore omitted.

- Not controlling for the effect of the specific texts as not enough levels (6) for random effects and when introduced as level 2 predictor, too conservative (the effect of the emotional differences between the texts would be removed).

### Choosing link family:
link-distributions (usually only minor impact on results):
logit = logistic,
probit = gaussian,
cloglog = extreme value distribution
http://bayesium.com/which-link-function-logit-probit-or-cloglog/

The choice should be made based on some combination of:
- Knowledge of the response distribution,
- Theoretical considerations, and
- Empirical fit to the data.
https://stats.stackexchange.com/questions/20523/difference-between-logit-and-probit-models

### interpretation of summary of fitted model
- Estimate is the mean of the posterior distribution, and corresponds to the frequentist point estimate
- Est.Error is the standard deviation of the posterior distribution
- thresholds in ordinal models are called "intercepts"
marginal effects for ordinal models: https://github.com/paul-buerkner/brms/issues/190

### posterior predictive checks
for ordinal models pp_check not adequate: https://github.com/stan-dev/bayesplot/issues/73
ppc: https://mc-stan.org/bayesplot/articles/graphical-ppcs.html

## load data
```{r }
# load data
df <- read_csv("df_TEEM_final.csv")

# rename some variables
df <- df %>% dplyr::rename("participant" = "subject_nr", "text" = "text_pic", "valence_post" = "SAM_LIKERT_POST")
```


## standardize predictors (helps for model convergence and for the interpretation of the interaction effects)
```{r}
df <- df %>% 
  mutate(
    mean_interest = scale(mean_interest, center = T, scale = T),
    mean_boredom = scale(mean_boredom, center = T, scale = T),
    mean_valence = scale(mean_valence, center = T), scale = T)

df <- df %>% 
  mutate(
    sd_interest = scale(sd_interest, center = T, scale = T),
    sd_boredom = scale(sd_boredom, center = T, scale = T),
    sd_valence = scale(sd_valence, center = T), scale = T)

df <- df %>% 
  mutate(
    peak10_interest = scale(peak10_interest, center = T, scale = T),
    peak10_boredom = scale(peak10_boredom, center = T, scale = T),
    peak10_valence_pos = scale(peak10_valence_pos, center = T, scale = T),
    peak10_valence_neg = scale(peak10_valence_neg), center = T, scale = T)
```


# 1. INTEREST
## INTEREST restricted model
```{r}
# complete cases only (drop NAs)
dfsub <- df %>% dplyr::select(participant, interested_post, mean_interest, sd_interest, peak10_interest) %>% drop_na()

# restricted model
m0i_cloglog <- brm(
          interested_post ~ 1 + (1|participant),
          family = cumulative("cloglog"),                    
          prior = prior(cauchy(0, 10), class = sd), 
          iter = 4000, warmup = 2000, chains = 4, cores = 4,
          inits = 0,
          data = dfsub,
          save_all_pars = T) # needed for bayes factor later on

summary(m0i_cloglog)

# other link functions
m0i_logit <-  update(m0i_cloglog,
                    family = cumulative("logit"))

m0i_probit <-  update(m0i_cloglog,
                      family = cumulative("probit"))
# compare different link functions using assimilation of leave-one-out-cross validation (looic)
m0i_logit <- add_criterion(m0i_logit,"loo")
m0i_probit <- add_criterion(m0i_probit,"loo", reloo = TRUE) #"reloo = TRUE" actually calculates MCMC for problematic observations 
m0i_cloglog <- add_criterion(m0i_cloglog ,"loo", reloo = TRUE)
print(loo_compare(m0i_logit, m0i_probit, m0i_cloglog, criterion="loo"), simplify = F)  # cloglog 1.5-3.5 SD better

# chosen F distribution (Link function), final Null-Modell
m0i <- m0i_cloglog
```

## INTEREST mean
```{r}
# full model including FaceReader's estimate as predictor
m1_imean <-  update(m0i, formula. = ~ . + mean_interest,
                    prior = c(prior(normal(0, 10), class = b), 
                              prior(cauchy(0, 10), class = sd)), 
                    newdata = dfsub,
                    save_all_pars = T, 
                    sample_prior = TRUE)

## model parameter
summary(m1_imean)

# plots
plot(marginal_effects(m1_imean, "mean_interest", method = "fitted", categorical = F), points = T) # shows the strong influence of two obervations
  
# Model without "outliers"
dfsub_out <- dfsub %>% 
  filter(mean_interest < (mean(mean_interest) + 4*sd(mean_interest)) & mean_interest > (mean(mean_interest) - 4*sd(mean_interest))) 
dfsub_out # two observations less

dfsub %>% filter(mean_interest > (mean(mean_interest) + 4*sd(mean_interest)) | mean_interest < (mean(mean_interest) - 4*sd(mean_interest))) # ouliers are from one participant (highly expressive in video)


m1_imean_out <-  update(m0i, formula. = ~ . + mean_interest,
                    prior = c(prior(normal(0, 10), class = b), 
                              prior(cauchy(0, 10), class = sd)), 
                    newdata = dfsub_out,
                    save_all_pars = T, 
                    sample_prior = TRUE)
summary(m1_imean_out)

## plots    
plot(marginal_effects(m1_imean_out, "mean_interest", method = "fitted", categorical = F), points = T)

# chose final model
m1_imean <- m1_imean_out
```



## INTEREST SD * mean
```{r}
# the participant from before also outlier in SD? --> No! --> leave it in
dfsub_out %>% filter(sd_interest > (mean(sd_interest) + 4*sd(sd_interest)) | sd_interest < (mean(sd_interest) - 4*sd(mean_interest))) # ouliers are from one participant (highly expressive in video)


# full model including FaceReader's estimates as predictor
m1_imeanxsd <-  update(m0i, formula. = ~ . + mean_interest*sd_interest,
                    prior = c(prior(normal(0, 10), class = b), 
                              prior(cauchy(0, 10), class = sd)),
                    control = list(adapt_delta = 0.99),
                    newdata = dfsub_out, # without outliers of mean interest
                    save_all_pars = T)

## model indicator
summary(m1_imeanxsd)

# plots
plot(marginal_effects(m1_imeanxsd,"mean_interest:sd_interest"), points = T, point_args = list(width = 0.9, alpha = 0.6)) 
plot(marginal_effects(m1_imeanxsd,"mean_interest"), points = T, point_args = list(width = 0.9, alpha = 0.6)) 
plot(marginal_effects(m1_imeanxsd,"sd_interest"), points = T, point_args = list(width = 0.9, alpha = 0.6)) 
```

## INTEREST mean of peaks
```{r}
#remove outliers from mean also in peak:
dfsub_outpeak <- dfsub %>% filter(peak10_interest < (mean(peak10_interest) + 4*sd(peak10_interest)) & peak10_interest > (mean(peak10_interest) - 4*sd(peak10_interest))) 
dfsub_outpeak

dfsub %>% filter(peak10_interest > (mean(peak10_interest) + 4*sd(peak10_interest)) | peak10_interest < (mean(peak10_interest) - 4*sd(peak10_interest)))  # 2 ouliers are from the same participant as before

# model including outliers
m1_ipeak <- update(m0i, formula. = ~ . + peak10_interest,
                    prior = c(prior(normal(0, 10), class = b), 
                              prior(cauchy(0, 10), class = sd)),
                              newdata = dfsub,
                              save_all_pars = T)
summary(m1_ipeak)

# model without outliers
m1_ipeak_out <- update(m0i, formula. = ~ . + peak10_interest,
                    prior = c(prior(normal(0, 10), class = b), 
                              prior(cauchy(0, 10), class = sd)),
                              newdata = dfsub_outpeak,
                              control = list(adapt_delta = 0.99),
                              save_all_pars = T)

## model parameter 
summary(m1_ipeak_out)

# plots
plot(marginal_effects(m1_ipeak,"peak10_interest", categorical = F), points = T) 
plot(marginal_effects(m1_ipeak_out,"peak10_interest", categorical = F), points = T)

m1_peak <- m1_ipeak_out
```

# 2. BOREDOM
## BOREDOM restricted model

```{r}
# complete cases only (drop NAs)
dfsubb <- df %>% dplyr::select(participant, bored_post, mean_boredom, sd_boredom, peak10_boredom) %>% drop_na()

# restricted model
m0b_cloglog <- brm(
          bored_post ~ 1 + (1|participant),
          family = cumulative("cloglog"),
          prior = prior(cauchy(0, 10), class = sd),
          iter = 4000, warmup = 2000, chains = 4, cores = 4,
          inits = 0,
          data = dfsubb,
          save_all_pars = T)

summary(m0b_cloglog)

# other link functions
m0b_logit <-  update(m0b_cloglog,
                    family = cumulative("logit"))

m0b_probit <-  update(m0b_cloglog,
                      family = cumulative("probit"))

# compare different link functions
m0b_logit <- add_criterion(m0b_logit,"loo")
m0b_probit <- add_criterion(m0b_probit,"loo")
m0b_cloglog <- add_criterion(m0b_cloglog ,"loo")
print(loo_compare(m0b_logit, m0b_probit, m0b_cloglog, criterion="loo"), simplify = F)

# chosen F distribution (link function)
m0b <- m0b_cloglog
```

## BOREDOM mean

```{r}
# full model including FaceReader's estimate as predictor
m1_bmean <-  update(m0b, formula. = ~ . + mean_boredom,
                    prior = c(prior(normal(0, 10), class = b), 
                              prior(cauchy(0, 10), class = sd)), 
                    newdata = dfsubb,
                    save_all_pars = T)

## model parameter
summary(m1_bmean)

# plots
plot(marginal_effects(m1_bmean, "mean_boredom", categorical = F), points = T) 
```

## BOREDOM mean and SD
```{r}
# full model including FaceReader's estimates as predictor
m1_bmeanxsd <-  update(m0b, formula. = ~ . + mean_boredom*sd_boredom,
                    prior = c(prior(normal(0, 10), class = b), 
                              prior(cauchy(0, 10), class = sd)),
                    newdata = dfsubb,
                    save_all_pars = T)

## model indicators 
summary(m1_bmeanxsd)

# plots
plot(marginal_effects(m1_bmeanxsd,"mean_boredom:sd_boredom", categorical = F), points = T)

## plot b densities and CIs
m1_bmeanxsd_ggs <- ggs(m1_bmeanxsd) # transforms the brms output into a longformat tibble (used to make different types of plots)
m1_bmeanxsd_ggs
ggplot(filter(m1_bmeanxsd_ggs, Parameter == "b_mean_boredom:sd_boredom", Iteration>1000), aes(x=value)) +
  geom_density(fill = "orange", alpha = .5) + geom_vline(xintercept = 0, col="red", size=1) +
  scale_x_continuous(name="Value", limits=c(-1, 2)) + 
  labs(title="Posterior density of regression coefficient for mean of FaceReader's boredom") +
  theme_apa() + 
  geom_vline(xintercept = summary(m1_bmeanxsd)$fixed[7,3:4], col="blue", linetype=2) # 95% CrI

# 10-fold cross validation
m0b <- add_criterion(m0b, criterion =  "kfold", folds = "grouped", group = "participant")
m1_bmeanxsd <- add_criterion(m1_bmeanxsd, criterion = "kfold", folds = "grouped", group = "participant")
print(loo_compare(m0b, m1_bmeanxsd, criterion = "kfold"), simplify = T) ## Estimating out-of sample predictions (via 10-fold cross validation) of the interaction model, compared to a model with no predictors yielded better results for the model without the interaction. Accordingly, we consider this potential interaction effect as irrelevant. 
```

## BOREDOM mean of peaks
```{r}
m1_bpeak <-  update(m0b, formula. = ~ . + peak10_boredom,
                    prior = c(prior(normal(0, 10), class = b), 
                              prior(cauchy(0, 10), class = sd)),
                    newdata = dfsubb,
                    save_all_pars = T)

## model parameter 
summary(m1_bpeak)

# plots
plot(marginal_effects(m1_bpeak,"peak10_boredom", categorical = F), points = T) 
```

# 3. VALENCE
stronger priors for valence, as issues with convergence.
## VALENCE restricted model
```{r}
## select complete cases of relevant variables
dfsubv <- df %>% dplyr::select(participant, valence_post, mean_valence, sd_valence, peak10_valence_pos, peak10_valence_neg) %>% drop_na()

# restricted model
## probit-model
m0v_probit <- brm(
          valence_post ~ 1 + (1|participant),
          family = cumulative("probit"),
          prior = c(prior(normal(0, 1), class = Intercept), 
                    prior(cauchy(0, 1), class = sd)),
          iter = 4000, warmup = 2000, chains = 4, cores = 4,
          control = list(adapt_delta = 0.99, max_treedepth = 15),
          data = dfsubv,
          inits = 0,
          save_all_pars = T)
summary(m0v_probit)

# chosen F distribution (Link function)
m0v <- m0v_probit
```

## VALENCE mean
```{r}
# full model including FaceReader's estimate as predictor
m1_vmean <-  update(m0v, formula. = ~ . + mean_valence,
                    prior = c(prior(normal(0, 1), class = Intercept),
                              prior(normal(0, 1), class = b), 
                              prior(cauchy(0, 1), class = sd)), 
                    newdata = dfsubv,
                    save_all_pars = T, 
                    sample_prior = T)

## model parameter
summary(m1_vmean)

# plots
plot(marginal_effects(m1_vmean, "mean_valence", categorical = F), points = T, point_args = c(alpha = 0.8)) 
```

## VALENCE mean and SD
```{r}
# full model including FaceReader's estimates as predictor
m1_vmeanxsd <-  update(m0v, formula. = ~ . + mean_valence*sd_valence,
                    prior = c(prior(normal(0, 1), class = Intercept),
                              prior(normal(0, 1), class = b), 
                              prior(cauchy(0, 1), class = sd)),
                    control = list(adapt_delta = 0.99, max_treedepth = 15),
                    newdata = dfsubv,
                    save_all_pars = T,
                    sample_prior = T)

## model parameter
summary(m1_vmeanxsd)

# plots
plot(marginal_effects(m1_vmeanxsd,"mean_valence:sd_valence"), points = T) 
```


## VALENCE mean of peaks
```{r}
m1_vpeak <-  update(m0v, formula. = ~ . + peak10_valence_neg + peak10_valence_pos,
                    prior = c(prior(normal(0, 1), class = Intercept),
                              prior(normal(0, 1), class = b), 
                              prior(cauchy(0, 1), class = sd)),
                    newdata = dfsubv,
                    save_all_pars = T)

## model parameter 
summary(m1_vpeak)

# plots
plot(marginal_effects(m1_vpeak,"peak10_valence_neg", categorical = F), points = T) 
plot(marginal_effects(m1_vpeak,"peak10_valence_pos", categorical = F), points = T) 
```

